---
title: 'Causal Inference on Emotions in Conversations'
author: 'Huy Tran'
output: html_document
---

```{r include = FALSE}
knitr::opts_chunk$set(include = FALSE, echo = FALSE, message = TRUE, warning = TRUE)
```

## Causal Inference Algorithm Installation

### pcalg

- Install the 2 dependencies from bioconductor
<http://www.bioconductor.org/install/>

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(c("graph", "RBGL", "Rgraphviz"))
```

- Install the main library
<https://cran.r-project.org/web/packages/pcalg/>

```{r}
install.packages('pcalg')
```

### Invariant Causal Prediction

- Install the package
<https://cran.r-project.org/web/packages/InvariantCausalPrediction/>
```{r}
install.packages('InvariantCausalPrediction')
```


## Load algorithms

```{r}
library('pcalg')
library('InvariantCausalPrediction')
```

## Variables of Interest

Emotion: labelled within the dataset
Has_Dialogue_Acts: tagged by a ISO-standard tagger
Has_Entities: classified by spacy
Speaker_ID: recoded from Speaker
Listener_ID: recoded from Speaker
Season: labelled within the dataset

### Casting to discrete values starting from 0 in order to apply algorithms

```{r}
emo <- read.csv('datasets/meld_yolo.csv')

data <- subset(emo, select = c('Emotion', 'Has_Entities', 'Dialogue_Act_Comm_Func', 'Speaker_ID', 'Listener_ID', 'Season'))

data$Emotion <- as.numeric(data$Emotion) - 1
data$Has_Entities <- as.numeric(data$Has_Entities) - 1
data$Dialogue_Act_Comm_Func <- as.numeric(data$Dialogue_Act_Comm_Func) - 1
data$Speaker_ID <- as.numeric(data$Speaker_ID)
data$Listener_ID <- as.numeric(data$Listener_ID)
data$Season <- as.numeric(data$Season) - 1
```

## Apply algorithms

### 1) InvariantCausal

```{r}
icp.data <- subset(data, select = c('Speaker_ID', 'Listener_ID', 'Has_Entities', 'Dialogue_Act_Comm_Func'))
icp.fit <- InvariantCausalPrediction::ICP(as.matrix(icp.data), data[['Emotion']], data[['Season']])
hidden.icp.fit <- InvariantCausalPrediction::hiddenICP(as.matrix(icp.data), data[['Emotion']], data[['Season']])

icp.fit
hidden.icp.fit
```


### 2) PC

```{r}
V <- colnames(data)
suffStat <- list(dm = as.matrix(data), adaptDF = FALSE)
pc.D <- pc(
  suffStat,
  ## independence test: G^2 statistic
  indepTest = disCItest, alpha = 0.01, labels = V, verbose = FALSE
)
summary(pc.D)
```


### 3) FCI

```{r}
V <- colnames(data)
suffStat <- list(dm = as.matrix(data), adaptDF = FALSE)
fci.D <- fci(
  suffStat,
  ## independence test: G^2 statistic
  indepTest = disCItest, alpha = 0.01, labels = V, verbose = FALSE
)
summary(fci.D)
```
